---
title: "Demo of latent Tree Aggregation Surrogate Variable Analysis"
author: "Christophe Ambroise"
date: "24/11/2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(LITree)
```


# Data Simulation

To simulate Gaussian data with given covariance structure and missing variables:

- chose a graph type (erdös, tree, star) 
- deduce a covariance matrix via the construction  of a GGM.model object
- draw a random sample from the GGM.model object via the `randomSample()` method
```{r graph-model}
erdos.graph <- graph.model$new(type = "erdos",size=30, p.or.m = 0.2)
plot(erdos.graph,edge.arrow.size=0, vertex.color="gold", 
     vertex.size=15, vertex.frame.color="gray", 
     vertex.label.color="black", 
     vertex.label.cex=0.8,
     vertex.label.dist=0)

star.graph <- graph.model$new(type = "starerdos",size=30, p.or.m = 0.05)
plot(star.graph,edge.arrow.size=0, vertex.color="gold", 
     vertex.size=15, vertex.frame.color="gray", 
     vertex.label.color="black", 
     vertex.label.cex=0.8,
     vertex.label.dist=0)
```





```{r GGM-simul}
star.model <- GGM.model$new(graph=star.graph) 
plot(star.model)
star.model$randomSample(n=50)  # create a random sample asscesible as public variable X in the star.model
```


# Graph inference


To infer the graphical model the class GGM.fit uses a simple data table (assumed to be scaled).
If the ground truth is know (for example when simulated data is available), precision recall and roc curve
can be computed using the GGM.evaluation class.

## Testing Glasso with no missing variable

```{r GGM.fit}
star.model.fit <-GGM.fit$new(star.model$getX(),method="glasso")
str(star.model.fit)
testingGlasso<-GGM.experiment$new(X.list = list(star.model$getX()), adjmat = star.model$getAdjmat())
testingGlasso$run()
testingGlasso$auc()
testingGlasso$roc.plot()
```

## Comparing Chow-Liu and Glasso with no missing variable

```{r }
testingGlasso<-GGM.experiment$new(X.list = list(star.model$getX()), adjmat = star.model$getAdjmat(),methods=c("glasso","chow.liu"))
testingGlasso$run()
testingGlasso$roc.plot()
print(auc<-testingGlasso$auc())
```
## Comparing Chow-Liu and Glasso with no missing variable using 10 samples of size 30

```{r }
X.list<- lapply(vector(mode = "list", length = 10), function(x) {star.model$randomSample(30)
                                                                 star.model$getX()  } )
testingGlasso<-GGM.experiment$new(X.list = X.list, adjmat = star.model$getAdjmat(),methods=c("glasso","chow.liu"))
testingGlasso$run()
testingGlasso$roc.plot()
print(auc<-testingGlasso$auc())
```

## Comparing Chow-Liu and Glasso with no missing variable using bagging

```{r }
testingGlasso<-GGM.experiment$new(X.list = list(star.model$getX()), adjmat = star.model$getAdjmat(),methods=c("glasso","chow.liu"))
testingGlasso$run(bagging=TRUE)
testingGlasso$roc.plot()
print(auc<-testingGlasso$auc())
```


# Introduction missing variable (latent roots)
The GGM.model class allows to specify a number of missing variables, which are chosen among the nodes of highest degree.
If there are some missing variable two matrices of adjacency are available:

- the full adjacency matrix
- the conditional adjacency matrix (relative to the observed variables)

And after the simulation of a random sample to data matrices are available: 

- The full data matrix $X$, with observed and latent variables
- The observed data  matrix $X_{obs}$ 

When the model is fitted with observed data, two evaluation can be performed, if the algorithm takes 
into account missing data

- the evaluation on the conditional graph, 
- the evaluation on the full graph.

# Testing Glasso with one missing variable
```{r GGM.model.with.missing.var}
star.model.missing <- GGM.model$new(graph=star.graph,nb.missing.var= 1) 
dim(star.model.missing$getAdjmat())
dim(star.model.missing$getAdjmatCond())
star.model.missing$randomSample(n=60)
dim(star.model.missing$getX())
dim(star.model.missing$getXobs())

star.model.missing.fit <-GGM.fit$new(star.model.missing$getXobs(),fit.number = 20,method="glasso")
star.model.missing.test <- GGM.evaluation$new(star.model.missing.fit, star.model.missing,type="conditional" )
star.model.missing.test$roc.plot()
star.model.missing.test$pr.plot()
star.model.missing.test$auc()
```


## Testing em Glasso for infering the full graph
```{r}
star.model.missing.fit.em <-GGM.fit$new(star.model.missing$getXobs(),method = "em.glasso",nb.missing.var = 1)
star.model.missing.test.full <- GGM.evaluation$new(star.model.missing.fit.em, star.model.missing,type="full" )
star.model.missing.test.full$roc.plot()
star.model.missing.test.full$pr.plot()
star.model.missing.test.full$auc()
```

## Testing EM latent roots algorithm for infering the full graph
```{r}
star.model.missing.fit.em <-GGM.fit$new(star.model.missing$getXobs(),method = "em.latent.roots",nb.missing.var = 1)
star.model.missing.test.full <- GGM.evaluation$new(star.model.missing.fit.em, star.model.missing,type="full" )
star.model.missing.test.full$roc.plot()
star.model.missing.test.full$pr.plot()
star.model.missing.test.full$auc()
```
## Testing EM latent roots algorithm for infering the conditional graph
```{r}
star.model.missing.fit.em <-GGM.fit$new(star.model.missing$getXobs(),method = "em.latent.roots",nb.missing.var = 1)
star.model.missing.test.full <- GGM.evaluation$new(star.model.missing.fit.em, star.model.missing,type="conditional" )
star.model.missing.test.full$roc.plot()
star.model.missing.test.full$pr.plot()
star.model.missing.test.full$auc()
```

## Testing em Glasso for infering the conditional graph
```{r}
star.model.missing.fit.em <-GGM.fit$new(star.model.missing$getXobs(),method = "em.glasso",nb.missing.var = 1)
star.model.missing.test.full <- GGM.evaluation$new(star.model.missing.fit.em, star.model.missing,type="conditional" )
star.model.missing.test.full$roc.plot()
star.model.missing.test.full$pr.plot()
star.model.missing.test.full$auc()
```
# Testing Experiment
### With one sample verus 30 boostraped samples or 3 cv
```{r}
library(latentRootsGGM)
star.model<-GGM.model$new(type = "erdos",size=50, p.or.m = 0.05)
#X.list<- lapply(vector(mode = "list", length = 1), function(x){ x<-star.model$randomSample(n=30)})
X<-star.model$randomSample(n=15)
X.list<-list(X)
test<- GGM.experiment$new(X.list,methods=c("glasso","chow.liu"),adjmat=star.model$getAdjmat())
test$run()
test$roc.plot()

X.list<- lapply(vector(mode = "list", length = 30), function(elt) elt<- X[sample(1:nrow(X),nrow(X),replace=TRUE),] )
test<- GGM.experiment$new(X.list,methods=c("glasso","chow.liu"),adjmat=star.model$getAdjmat())
test$run()
test$roc.plot()

sample.cv<-sample(1:2,nrow(X),replace=TRUE)
X.list<- Map(function(x,i)  X[sample.cv==i,], vector(mode = "list", length = max(sample.cv)), sample.cv )
test<- GGM.experiment$new(X.list,methods=c("glasso","chow.liu"),adjmat=star.model$getAdjmat())
test$run()
test$roc.plot()

```
### With 10 bootstrapped samples
```{r}
library(latentRootsGGM)
erdos.model<-GGM.model$new(type = "starerdos",size=35, p.or.m = 0.05)
erdos.model$randomSample(40)
erdos.model.fit <-GGM.fit$new(erdos.model$getX(),method="glasso")
erdos.model.fit$run()
testingGlasso<-GGM.evaluation$new(erdos.model.fit,erdos.model)
testingGlasso$roc.plot()
erdos.model.fit$run(bagging=TRUE)
testingGlasso<-GGM.evaluation$new(erdos.model.fit,erdos.model)
testingGlasso$roc.plot()
erdos.model.fit <-GGM.fit$new(erdos.model$getX(),method="chow.liu")
erdos.model.fit$run(bagging=TRUE)
testingGlasso<-GGM.evaluation$new(erdos.model.fit,erdos.model)
testingGlasso$roc.plot()
```

# Test of experiment
```{r}
star.model<-GGM.model$new(type = "erdos",size=50, p.or.m = 0.05)
X.list<-X.list<- lapply(vector(mode = "list", length = 10), function(x){ x<-star.model$randomSample(n=30)})
test<- GGM.experiment$new(X.list,methods=c("glasso","chow.liu"),adjmat=star.model$getAdjmat())
test$run()
test$roc.plot()
```

# Test of experiment with missing variable
```{r}
star.model<-GGM.model$new(type = "erdos",size=50, p.or.m = 0.05,nb.missing.var=1)
X.list<-X.list<- lapply(vector(mode = "list", length = 10), function(x){ x<-star.model$randomSample(n=30)})
test<- GGM.experiment$new(X.list,methods=c("glasso","chow.liu"),adjmat=star.model$getAdjmat())
test$run()
test$roc.plot()
```
```{r}
star.model<-GGM.model$new(type = "starerdos",size=50, p.or.m = 0.05,nb.missing.var=1)
dim(star.model$getAdjmat())
dim(star.model$getAdjmatCond())
star.model$randomSample(50)
fit.em.latent.roots<- GGM.fit$new(X=star.model$getXobs(), nb.missing.var = 1, method="em.latent.roots")
```


